---
output:
  html_document:
    code_folding: hide
    theme: journal
  pdf_document: default
authors: Dejia Tang, Zicong Wei
---
# **Can Comedy Be Horror:    **
# **Overlapping Genres in Films**
**Authors**: Dejia Tang, Zicong Wei   
**Last Edit**: `r format(Sys.time(), '%d %B, %Y')`   

```{r message=FALSE, echo=FALSE, warning=FALSE}
library(mdsr)
library(RMySQL)
library(tidyverse)
library(ggthemes)
```

```{r message=FALSE, echo=TRUE, warning=FALSE}
# Get dataframe using SQL
db <- dbConnect_scidb(dbname = "imdb")

# Tables are joined using ids to speed things up.
sql <- "
SELECT t.id, mi.info, t.production_year, it.info as type, mi.info_type_id
FROM title as t
LEFT JOIN movie_info as mi ON mi.movie_id = t.id
LEFT JOIN info_type as it ON mi.info_type_id = it.id
WHERE production_year is not NULL;
"
df <- db %>%
  dbGetQuery(sql) %>%
  filter(production_year < 2017)

# Get dataframe using SQL
df_country <- df %>%
  filter(info_type_id == 8) %>%
  mutate(country = info) %>%
  select(id, country, production_year)

df_genre <- df %>%
  filter(info_type_id == 3) %>%
  mutate(genre = info) %>%
  select(id, genre, production_year)

df_country_genre <- df_country %>%
  right_join(df_genre, on = id)


# First Chart: genre variety
df_usa <- df_country_genre %>%
  filter(country == 'USA')

# Calculate totol number of geners appeared each year.
df1_total <- df_usa %>%
  group_by(production_year) %>%
  summarise(genre_total = n_distinct(genre))

# Calculate average number of geners per movie in each year.
df1_avg <- df_usa %>%
  group_by(production_year, id) %>%
  summarise(genre_count = n_distinct(genre)) %>%
  group_by(production_year) %>%
  summarise(genre_avg = mean(genre_count))

# Join the two tables together.
df1 <- df1_total %>%
  left_join(df1_avg, on = 'production_year') %>%
  gather(key = "value_type", value = "value",-production_year)

# The plot.
plot1 <- df1 %>%
  ggplot(aes(production_year, value)) +
  geom_line() +
  facet_wrap(~factor(value_type, labels = c("Average Number of Genres Per Film in the U.S.", "Total Number of Genres in the U.S.")), scale = "free") +
  ggtitle("Figure 1: Genre Counts of American Films Over time")

# Themes
plot1 <- plot1 +
  theme_fivethirtyeight() +
  theme(panel.grid.major = element_blank(), axis.title.x=element_blank(), axis.title.y=element_blank())

# Second Chart: Change of Popularity for Short, Drama and Comedy
# Calculate totol number of movies appeared each year.
df2_total <- df_usa %>%
  group_by(production_year) %>%
  summarise(movie_total = n_distinct(id))

# Join and calculate totol each gener's proportion for each year.
df2 <- df_usa %>%
  left_join(df2_total, on = production_year) %>%
  group_by(production_year, genre) %>%
  summarise(proportion = n()/first(movie_total))

# Create the plot.
plot2 <- df2 %>%
  filter(genre == 'Short' | genre == 'Comedy' | genre == 'Musical') %>%
  ggplot(aes(production_year, proportion, col = genre)) +
  geom_line() +
  ggtitle("Figure 2: Proportion of Short, Musical, and Comedy in the U.S.") +
  theme_fivethirtyeight() +
  theme(panel.grid.major = element_blank(),
        axis.title.x=element_blank(), axis.title.y=element_blank()) +
  geom_segment(aes(x = 1947, y =0, xend = 1947, yend = 0.5), 
               color = "black", size =0.1) +
  geom_text(aes(x = 1947, y = 0.55, label = "United States v.\nParamount Pictures, Inc. 1947"), 
            size = 3, color = "black") +
  geom_segment(aes(x = 1994, y =0, xend = 1994, yend = 0.25), 
               color = "black", size =0.1) +
  geom_text(aes(x = 1994, y = 0.3, label = "Independent\nFilms Rise 1994"), 
            size = 3, color = "black") +
  geom_segment(aes(x = 2000, y =0, xend = 2000, yend = 0.43), 
               color = "black", size =0.1) +
  geom_text(aes(x = 2000, y = 0.46, label = "Digital Camera 2000"), 
            size = 3, color = "black") +
  geom_segment(aes(x = 2009, y =0, xend = 2009, yend = 0.55), 
               color = "black", size =0.1) +
  geom_text(aes(x = 2009, y = 0.58, label = "3D Avatar 2009"), 
            size = 3, color = "black") +
  geom_segment(aes(x = 1976, y =0, xend = 1976, yend = 0.35), 
               color = "black", size =0.1) +
  geom_text(aes(x = 1976, y = 0.4, label = "Star Wars\n'blockbuster' 1976"), 
            size = 3, color = "black") +
  geom_segment(aes(x = 1927, y =0, xend = 1927, yend = 0.7), 
               color = "black", size =0.1) +
  geom_text(aes(x = 1927, y = 0.8, label = "First Film with Sound"), 
            size = 3, color = "black")

# Third Chart: Composition of Comedies
short <- df_usa %>%
  filter(genre == 'Comedy' & production_year > 1927) %>%
  group_by(id) %>%
  summarise(production_year = first(production_year))

other_genres <- df_usa %>%
  filter((genre == 'Horror' | genre == 'Romance' | genre == 'Family' | genre == 'War') & production_year > 1927) %>%
  group_by(id) %>%
  summarise(production_year = first(production_year))

df3 <- df_usa %>%
  #filter(id %in% comedies$id) %>%
  #filter(genre != 'Comedy') %>%
  filter(id %in% short$id & id %in% other_genres$id) %>%
  filter(genre == 'Horror' | genre == 'Romance' | genre == 'Family' | genre == 'War') %>%
  group_by(production_year, genre) %>%
  summarize(count = n())

plot3 <- df3 %>%
  ggplot(aes(x = production_year, y = count, fill= genre)) +
  geom_bar(stat="identity", position='fill', alpha=0.8) +
  theme_fivethirtyeight() +
  scale_fill_manual(
    values = c("#f7c143", "#556eed", "#f2763c", "#a5a5a5", "#9B002E", "#222222"),
    name = "Film Genre") +
  ggtitle("Figure 4: Genre Composition of Comedies Over Time in the U.S.")



# Fourth Chart: Composition of Short Films
short <- df_usa %>%
  filter(genre == 'Short' & production_year > 1927) %>%
  group_by(id) %>%
  summarise(production_year = first(production_year))

other_genres <- df_usa %>%
  filter((genre == 'Sci-Fi' | genre == 'Romance' |genre == 'Fantasy' | genre == 'Adventure'| genre == 'Animation') & production_year > 1927) %>%
  group_by(id) %>%
  summarise(production_year = first(production_year))

df4 <- df_usa %>%
  filter(id %in% short$id & id %in% other_genres$id) %>%
  filter(genre == 'Sci-Fi' | genre == 'Romance' | genre == 'Fantasy' | genre == 'Adventure'| genre == 'Animation') %>%
  group_by(production_year, genre) %>%
  summarize(count = n())

plot4 <- df4 %>%
  ggplot(aes(x = production_year, y = count, fill= genre)) +
  geom_bar(stat="identity", position='fill', alpha=0.8) +
  theme_fivethirtyeight() +
  scale_fill_manual(
    values = c("#f7c143", "#556eed", "#f2763c", "#a5a5a5", "#9B002E", "#222222"),
    name = "Film Genre") +
  ggtitle("Figure 3: Genre Composition of Short Films Over Time in the U.S.")


# Fifth Chart: Composition of Musicals
musical <- df_usa %>%
  filter(genre == 'Musical' & production_year > 1927) %>%
  group_by(id) %>%
  summarise(production_year = first(production_year))

other_genres <- df_usa %>%
  filter((genre == 'Romance' | genre == 'Comedy' | genre == 'Western' | genre == 'Animation' | genre == 'Drama')) %>%
  group_by(id) %>%
  summarise(production_year = first(production_year))

df5 <- df_usa %>%
  filter(id %in% musical$id & id %in% other_genres$id) %>%
  filter(genre == 'Romance'  | genre == 'Western' | genre == 'Animation' | genre == 'Comedy' | genre == 'Drama') %>%
  group_by(production_year, genre) %>%
  summarize(count = n())

plot5 <- df5 %>%
  ggplot(aes(x = production_year, y = count, fill= genre)) +
  geom_bar(stat="identity", position='fill', alpha=0.8) +
  theme_fivethirtyeight()  +
  scale_fill_manual(
    values = c("#f7c143", "#556eed", "#f2763c", "#a5a5a5", "#9B002E", "#222222"),
    name = "Film Genre") +
  ggtitle("Figure 5: Genre Composition of Musicals Over Time in the U.S.")
```

While Hollywood is making movies with seemingly repetitive themes, the invention of digital cameras also allows independent filmmakers to experiment with new genres. Is the number of genres expanding? Do individual genres remain the same meaning over time? How do genres interact with each other?      

Film genre is a blueprint of the industry and a consensus of the society. Producers and audiences envision films' content based on a conventional framework of genre films, but these genres also overlap with each other, providing more diverse features. It is noteworthy that in a certain period, people favor specific genres’ combination, and this popularity of specific genres reflects people’s social attitude and interests. Shorts, Comedies, and Musicals are important categories reflect the genres’ changing trend throughout film history. Their various overlap with other genres represents the change in the U.S. society, such as technical innovation and social improvement.     

## Overall Trend
```{r message=FALSE, echo=TRUE, warning=FALSE, fig.align="center", fig.width=10, fig.height=6.5}
plot1
```

A film can fit more than one genre. The number of overlapping genres varies throughout film history. Using the IMDB dataset, we examine the change of genre overlapping since 1980 when the film was created[^1].  

The total number of genres increased and became fixed around 28 in the 1950s. More importantly, the average number of film genres employed in a single film has been changed significantly over decades.  

The first plot shows that this number reached its peak around the 1940s and dramatically decreased after that period. It increased greatly again in the 2000s. There should be some events at that time to explain the drastic changes.  

Also, the increase from 2006 indicates that movies made today tend to have overlapping genres that are different from the past. Old movies tended to fit squarely in one genre, while more recent movies tend to span multiple genres.

```{r message=FALSE, echo=TRUE, warning=FALSE, fig.align="center", fig.width=10, fig.height=6.5}
plot2
```
Plot2 presents the proportion of films' genres in each year. Some genres have an important influence on the genre overlapping.      

Comedy, Musical, and Short are three influential genres in IMDB dataset. From this graph, we can see the trend of these genres.         

## Short Films
```{r message=FALSE, echo=TRUE, warning=FALSE, fig.align="center", fig.width=10, fig.height=6.5}
plot4
```
Now let’s take a closer look at the genre “short film” and its composition. This genre rarely comes to people’s minds when they are asked about what kinds of films they like. This may be caused by the fact that the label “short film” is not describing a film’s content or style, but its format: the running time should be less than 40mins. Short films are often screened at film festivals and made by independent filmmakers for nonprofit and with a low budget. They also frequently serve as samples that independent filmmakers produce to gain notice and fundings from investors and studios.      

Figure 3 present five major overlapping genres of short films including adventure, animation, fantasy, romance and science fiction. The pattern is almost perceivable at first glance: short animations dominated the category for a long time in history but gradually declined in the modern era. When short animation was in its golden age, almost all major film companies had specialized units to produce shorts[^2]. One milestone of short animation’s decline is the shutdown of [Warner Bros. Cartoons, Inc.](https://en.wikipedia.org/wiki/Warner_Bros._Cartoons), one of the biggest players of this field, in 1969.
      
One crucial reason for short animation’s abundance in history was low cost. However, as digital cameras became easily accessible for independent film producers, other subgenres inevitably expanded and took over: all other sub-genres here except animation have been increasing their shares through history.           

## Comedy
```{r message=FALSE, echo=TRUE, warning=FALSE, fig.align="center", fig.width=10, fig.height=6.5}
plot3
```
Comedy is one of the oldest genres of American cinema since the silent era. Figure 4 shows the six important genres overlapped with comedies and their changing popularities[^3].       

As the U.S. entered into World War II, military themes started to appear in comedies as well, such as topics of civil defense, boot-camp, and shore-leave. We can see that there’s a small bump of war subgenre around 1940 for this reason.      

Around 1950, television came into the media industry and was filled with family sitcoms. This can explain the large share of family comedy around then.      

Triggered by the success of [When Harry Met Sally](https://en.wikipedia.org/wiki/When_Harry_Met_Sally...) in 1989, romantic comedy started to win back its popularity, being indicated by a slight increase in romance comedies’ share.      

From plot 4, we can also observe the gain of popular for horror comedies starting around the 1970s. Some of them are of anti-war theme like [MASH](https://en.wikipedia.org/wiki/M*A*S*H) and [Catch 22](https://en.wikipedia.org/wiki/Catch-22).      

## Musical
```{r message=FALSE, echo=TRUE, warning=FALSE, fig.align="center", fig.width=10, fig.height=6.5}
plot5
```
n the golden age of the musical film, comedy, and romance contribute the most to its content.  

Around the early 1930s, in the first major musical “cycles,” the “backstage musical” was prevalent. In these films, song and dance are presented as performances and rehearsals of music shows.  

Later, in the mid-1930s, the “integrated musical” was popularized. Fred Astaire and Ginger Roger were the typical musical stars of this kind. The song and dance contribute to the plot as an expression of characters’ emotion.  

As the musical developed into the 1940s and 1950s, romance's proportion was larger. There was more genres’ courtship plot came to be associated with an underlying value conflict of the society at that time.  

Hollywood musicals, especially the Freed Unit, have a frivolous male character getting along with a serious female character, such as *Singin’ in the Rain* and *the Band Wagon*. The entertaining male character represents the popular culture, and the working female values represents the high culture.  

Hollywood as a mass culture was trying to challenge the imperative to work and the serious atmosphere. Musicals defend the pleasure of US popular cultures and various forms of entertainment[^4].  

After this classical period, musicals were no longer favored by the public, especially the old-fashion Hollywood musicals. Since the 1960s, lots of social movements impacted the society. Especially young people’s music tastes have changed, because of the influence of rock and roll and new popular cultures.  

During the 1980s to 1990s, animated films became popular with musicals, such as Disney's films *Beauty and the Beast* and *Lion King*. They all became blockbuster successes.   

In the 21st century, the musical film was reborn. Drama musicals and comedy-drama musicals such as Chicago and La La Land become more popular in musical genres. In recent years, animated musicals like Frozen and Coco also gained popularity among people. 

Conclusion & Counter-Argument

By analyzing the genre's coverage of films recorded in IMDB, we found that vast scale of social and technological changes may reshape the film industry’s genre composition. However, we also found that some trends seem to be immune to the changing waves. For example, the composition of the musical has not changed a lot across film history.
 
[^1]: GitHub Repository: https://github.com/zicongwei/Genre_Multiplicity_in_Film
[^2]: WikiPedia: [Short Films](https://en.wikipedia.org/wiki/Short_film)
[^3]: ["Comedy Films"](http://www.filmsite.org/comedyfilms2.html). Filmsite.org. Retrieved 2012-04-29.
[^4]: Thomas Schatz, “Film Genre and the Genre Film”
